---
title: "CrossLang"
output: html_document
---

Starting from database including features of demonstrative systems for wide language sample downloaded from <http://wals.info>
Cultural and geographical variables extracted from D-PLACE <http://d-place.info> and cross-linked via ISO and Glottolog code.

First Step: Import WALS data on number of demonstrative contrasts

``` {r DemsWALS, echo = FALSE}
pacman::p_load(lingtypology,readr, plotly, dplyr,lme4)


# Using features from WALS.
# Sadly the code has to be looked up on the website
demsWals <- wals.feature( features = c('41a'))
colnames(demsWals) <- c('Wals', 'NrContr', 'Latitude', 'Longitude', 'GlottologCode', 'Lang')

```

```{r DB, echo = FALSE, warning=FALSE}
library(readr)
library(lingtypology)

# Read in and preprocess linguistic data on additional features related to demonstratives
wals_extra <- wals.feature( features = c('37a', '43a', '45a'))
wals_extra <- subset(wals_extra, wals.code %in% unique(demsWals$Wals))
wals_extra <- wals_extra[, c(1, 3:5)]
colnames(wals_extra) <- c('Wals', 'DefArt', 'ThirdPersonProns', 'PolitenessDistinction')

```

Now let's read in the cultural data 
``` {r CultData, echo = FALSE, warning = FALSE}

# Read in and preprocess social data
socParams <- data.frame(read_csv('socParams.csv', skip = 1))
socParams <- socParams[, c(1,2,12,13,14,15,16,17,18,22,23,24,28,29,30,34,35,36,40,41,42)]
colnames(socParams) <- c('Source', 'SocName','GlottologCode', 'Lang', 'Wals', 'LangFam', 'ClassYear', 'ClassVal', 'ClassLegend', 'PopYear', 'PopVal', 'PopLegend', 'RigidityYear', 'RigidityVal', 'RigidityLegend', 'MeanSizeYear', 'MeanSizeVal', 'MeanSizeLegend', 'Class2Year', 'Class2Val', 'Class2Legend')
# Function from lingtypology to get ISO codes from Glottolog codes
iso_codes <- iso.gltc(unique(socParams$GlottologCode))
# Sto iso codes into database
iso_df <- data.frame(matrix(nrow = length(unique(socParams$GlottologCode)), ncol = 0))
iso_df$GlottologCode <- unique(socParams$GlottologCode)
iso_df$Wals <- iso_codes

# Fill in missing info on ISO codes
for (item in unique(socParams$GlottologCode)) {
  if (is.na(socParams$Wals[socParams$GlottologCode == item])){
  socParams$Wals[socParams$GlottologCode == item] <- iso_df$Wals[iso_df$GlottologCode == item]
  }
}

```


Same with environment data
```{r EnvData, warning = FALSE}

# Read in and preprocess environmental variables
envParams <- data.frame(read_csv('envParams.csv', skip = 1))
envParams <- envParams[, c(1,2,12,13,14,15,16,17,19,20,22,23)]
colnames(envParams) <- c('Source', 'SocName','GlottologCode', 'Lang', 'Wals', 'LangFam', 'AboveSeaVal', 'AboveSeaLegend', 'DistanceCoastVal', 'DistanceCoastLegend', 'MeanInclineVal', 'MeanInclineLegend')
# Do same trick to extract ISO codes
iso_codes <- iso.gltc(unique(envParams$GlottologCode))
iso_df <- data.frame(matrix(nrow = length(unique(envParams$GlottologCode)), ncol = 0))
iso_df$GlottologCode <- unique(envParams$GlottologCode)
iso_df$Wals <- iso_codes

# Fill in info too
for (item in unique(envParams$GlottologCode)) {
  if (is.na(envParams$Wals[envParams$GlottologCode == item])){
  envParams$Wals[envParams$GlottologCode == item] <- iso_df$Wals[iso_df$GlottologCode == item]
  }
}

```

Merging the dataframes
``` {r Merging, echo = FALSE}

demsWals <- merge(demsWals, envParams[, c(3,7:12)], by = c('GlottologCode'), all.x = TRUE)
demsWals <- merge(demsWals, socParams[, c(3,6, 7:21)], by = c('GlottologCode'), all.x = TRUE)
demsWals <- merge(demsWals, wals_extra, by = c('Wals'), all.x = TRUE)

# Making the variables more transparent
demsWals$Class2Val <- plyr::mapvalues(demsWals$Class2Val, from = unique(demsWals$Class2Val), to = c('No Distinction', 'Aristocracy', 'Wealth', 'Complex', 'Missing'))
demsWals$NrContr <- plyr::mapvalues(demsWals$NrContr, from = unique(demsWals$NrContr), to = c(2,3,4,1,5))

```



Plot data from both the databases
``` {r PlotData}
library(ggplot2)

# Let's see how old our data are
histYear <- ggplot(data = demsWals) + geom_histogram(aes(x = ClassYear), binwidth = 20) + xlab('') + ylab('Year')

# Distribution of distances from the coast, height and incline
histIncline <- ggplot(data = demsWals) + geom_histogram(aes(x = MeanInclineVal)) + xlab('') + ylab('Mean Incline (degrees)')
histSea <- ggplot(data = demsWals) + geom_histogram(aes(x = AboveSeaVal)) + xlab('') + ylab('Meters above sea level')
histHeight <- ggplot(data = demsWals) + geom_histogram(aes(x = DistanceCoastVal)) + xlab('') + ylab('Distance from coast (km)')

# Number of contrasts
nrCont <- ggplot(data = demsWals) + geom_bar(aes(x = NrCont)) + xlab('') + ylab('Count')

```

Now we have visualized the distribution of our variables. 
But what about the geographical distributions of our features?

``` {r PlotMaps}
library(plotly)

# Color styling
col <- c("green", "yellow", "red")

# geo styling
g <- list(
  showland = T,
  landcolor = toRGB("gray90"),
  subunitcolor = toRGB("gray90"),
  countrycolor = toRGB("white")
)

# Make interactive plot!
int_plot <- plot_geo(demsWals, lat = ~Latitude, lon = ~Longitude) %>%
  add_markers(text = ~paste(Lang, paste("Language: ", Lang),
                            paste("Latitude: ", round(Latitude, 2)), 
                            paste("Longitude: ", round(Longitude, 2)), 
                            paste("Nr Contrasts:", NrContr),
                            sep = "<br />"), color = ~NrContr, colors = col,  hoverinfo = "text")%>% add_annotations(text="# Contrasts", xref="paper", yref="paper",
                  x=1.02, xanchor="left",
                  y=0.8, yanchor="bottom",    # Same y as legend below
                  legendtitle=TRUE, showarrow=FALSE ) %>%
  layout(
    title = 'Demonstratives World Map', 
    xaxis = list(title = 'Longitude'), 
    yaxis = list(title='Latitude'), 
    geo = g,
    legend = list(x = 1.02, y = 0.8, xanchor= 'left', yanchor = "top"),
    autosize = F,
    width = 1600,
    height = 800
  )

```
We can look at how environmental and social features are distributed too!
Plotting this in logarithmic scale to see some differences
``` {r PlotEnv}

inclinePlot <- plot_geo(subset(demsWals, is.na(MeanInclineVal) == F), lat = ~Latitude, lon = ~Longitude) %>%
  add_markers(text = ~paste(Lang, paste("Language: ", Lang),
                            paste("Latitude: ", round(Latitude, 2)), 
                            paste("Longitude: ", round(Longitude, 2)), 
                            paste("Nr Contrasts:", NrContr),
                            sep = "<br />"), color = ~log(MeanInclineVal), hoverinfo = "text") %>% add_annotations(text="Incline ", xref="paper", yref="paper",
                  x=1.02, xanchor="left",
                  y=0.8, yanchor="bottom",    # Same y as legend below
                  legendtitle=TRUE, showarrow=FALSE ) %>%
  layout(
    title = 'Demonstratives World Map', 
    xaxis = list(title = 'Longitude'), 
    yaxis = list(title='Latitude'), 
    geo = g,
    legend = list(x = 1.02, y = 0.8, xanchor= 'left', yanchor = "top"),
    autosize = F,
    width = 1600,
    height = 800
  )


inclinePlot 
```

Social variables distribution
``` {r PopSizePlot}

popPlot <- plot_geo(subset(demsWals, is.na(PopVal) == F), lat = ~Latitude, lon = ~Longitude) %>%
  add_markers(text = ~paste(Lang, paste("Language: ", Lang),
                            paste("Latitude: ", round(Latitude, 2)), 
                            paste("Longitude: ", round(Longitude, 2)), 
                            paste("Nr Contrasts:", NrContr),
                            sep = "<br />"), color = ~log(PopVal), hoverinfo = "text") %>% add_annotations(text="Population Size ", xref="paper", yref="paper",
                  x=1.02, xanchor="left",
                  y=0.8, yanchor="bottom",    # Same y as legend below
                  legendtitle=TRUE, showarrow=FALSE ) %>%
  layout(
    title = 'Demonstratives World Map', 
    xaxis = list(title = 'Longitude'), 
    yaxis = list(title='Latitude'), 
    geo = g,
    legend = list(x = 1.02, y = 0.8, xanchor= 'left', yanchor = "top"),
    autosize = F,
    width = 1600,
    height = 800
  )

```

Run whatever toy analysis you wish
``` {r RunAnalyses}

library(dplyr)
library(lme4)

# Let's bin the data into two categories
demsWals$ContrBin[demsWals$NrContr <= 2] <- 1
demsWals$ContrBin[demsWals$NrContr >= 3] <- 2

# Toy hypothesis: more complexity is determined by more complex landscape
binMod <- glmer(data = demsWals, as.factor(ContrBin) ~ MeanInclineVal + (1|LangFam), family = 'binomial')
summary(binMod)

# Our model does not converge but whatever
# Table 
agg_data <- subset(demsWals, is.na(MeanInclineVal) == F) %>% group_by(as.factor(ContrBin)) %>% summarize(MeanIncline = mean(as.numeric(MeanInclineVal)))
head(agg_data)

```
